Spatial Reasoning and Statistical Graphics

### 2014-11-11 Paper/Draft1/TestingVisualAptitude.Rnw
mini edits during meeting

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
Computer Society, IEEEtran, journal, \LaTeX, paper, template.
\end{IEEEkeywords}}

@@

\comment{Next few steps for the analysis: \\
(1) I think we should add a linear model for lineup scores that includes the demographics of participants as well; \\
(2) could you do the scoring of lineups by type of plot used and investigate the relationship between those types of lineups and the other visual tests ... not sure that that will pay off, but it's interesting.\\
(3) in terms of why we would use lineups in a test comparison ... lineups give us a way to say 'this is correct' which regular graphics don't. So we should put a spin sentence on at the intro and the back that we are suggesting to use lineups as a tool to not only evaluate individuals (which we can see is possible) but also as a tool to evaluate charts ... which we've done before ... and tie it into the regular testing literature ...}

@@

<<VisReasoningSPM,echo=FALSE, include=T, eval=T, fig.width=8,fig.height=8, message=F, warning=F, fig.cap="Pairwise scatterplots of test scores. Lineup scores are most highly correlated with figure classification scores, and are also highly correlated with card rotation scores. Paper folding and card rotation scores are also highly correlated.\\label{fig:scatterplotmatrix}">>=
data <- ans.summary[,c("lineup", "vis_search", "fig_class", "card_rot", "folding")]
grid <- expand.grid(x=1:ncol(data), y=1:ncol(data))
grid <- subset(grid, x != y & x<y)

@@

ggplot()+
  geom_point(data=subset(all, points), aes(x=x, y=y)) + 
  geom_smooth(data=subset(all, points), aes(x=x, y=y), method="lm") + 
  geom_text(data=subset(all, !points), aes(x=x, y=y, label=paste0("Correlation = \n", label))) + 
  geom_line(data=densities, aes(x=x, y=y)) + 
  facet_grid(xvar~yvar, scales="free", labeller = labeller(xvar=fix.names, yvar=fix.names)) + 
  theme_bw() + 
  theme(axis.title=element_blank())
rm(list=c("data", "all", "densities"))
@

@@

\newtext{We see that again, PC1 accounts for general visual aptitude. PC2, however, emphasizes the relationship between figure classification and Lineup task 1 (and to a lesser extent, on lineup task 2). PC3 differentiates the lineup task 3 from the visual search task; other tasks do not impact PC3 to a meaningful degree. PC4 separates lineup test 2 from the paper folding and card rotation tests and emphasizes the overlap with the visual search test. PC5 separates lineup test 2 and the figure classification test from the overlap between the other lineup tests and the visual search test. PC 6 and 7 account for \Sexpr{as.numeric(round(100*importance[6], 1))}\% and \Sexpr{as.numeric(round(100*importance[7], 1))}\% of the variance, respectively; PC6 seems to differentiate the first two lineup tasks from the overlap between the third lineup task and the figure classification task.
}
<<lineupblockpca-biplot,echo=F, warning=F, message=F, fig.cap="This is awful, but at least shows everything", fig.env="figure*", fig.width=11, fig.height=10, out.width=".9\\linewidth">>=
data <- as.data.frame(lineup.block.pca$rotation)
grid <- expand.grid(x=1:7, y=1:7)
grid <- subset(grid, x != y)

  rbind.fill(
    data.frame(
      xvar = names(data)[xcol], 
      yvar = names(data)[ycol],
      x = 0,
      y = 0,
      xend = data[, xcol], 
      yend = data[, ycol], 
      label = rownames(data),
      points = T
    )
  )
}))
all$xvar <- factor(all$xvar, levels=names(data))
all$yvar <- factor(all$yvar, levels=names(data))

tmp <- rotation[,1:3]
names(tmp) <- c("xvar", "label", "rotation")
tmp$xvar <- tmp$yvar <- factor(tmp$xvar, levels=names(data))

all$label <- factor(all$label, 
                        levels=c("vis_search", "folding", "card_rot", "fig_class", "lineup_1", "lineup_2", "lineup_3"),
                        labels=c("Visual Search", "Paper Folding", "Card Rotation", "Figure Class.", "Lineup Test 1", "Lineup Test 2", "Lineup Test 3"))

ggplot()+
  geom_segment(data=subset(all, points), aes(x=x, y=y, xend=xend, yend=yend, colour=label), arrow=arrow(length=unit(.1, "cm"))) + 
#   geom_point(data=tmp, aes(x=0*rotation, y=rotation, colour=label)) + 
  geom_text(data=tmp, aes(x=ceiling(rotation)-.5, y=rotation, label=label, colour=label)) + 
  facet_grid(xvar~yvar) + 
  theme_bw() + 
  theme(axis.title=element_blank()) + 
  scale_colour_manual("Test", values = c("#f781bf", "#e41a1c", "#ff7f00", "#66a61e", "#1d91c0", "#225ea8", "#0c2c84"))
rm(list=c("data", "all"))
@
\subsection{Lineup Plot Types}

>>>>

% Note that keywords are not normally used for peerreview papers.
% \begin{IEEEkeywords}
% Computer Society, IEEEtran, journal, \LaTeX, paper, template.
% \end{IEEEkeywords}
}

@@

\comment{Next few steps for the analysis: \\
(1) I think we should add a linear model for lineup scores that includes the demographics of participants as well; \\
\textbf{Done; added to Results section}\\
(2) could you do the scoring of lineups by type of plot used and investigate the relationship between those types of lineups and the other visual tests ... not sure that that will pay off, but it's interesting.\\
\textbf{Done, along with an examination of each experiment separately}\\
(3) in terms of why we would use lineups in a test comparison ... lineups give us a way to say 'this is correct' which regular graphics don't. So we should put a spin sentence on at the intro and the back that we are suggesting to use lineups as a tool to not only evaluate individuals (which we can see is possible) but also as a tool to evaluate charts ... which we've done before ... and tie it into the regular testing literature ...\\
\textbf{A few sentences added, still need to tie into the existing literature}}

@@

\begin{figure}[htp]
<<VisReasoningSPM,echo=FALSE, include=T, eval=T, fig.width=8,fig.height=8, message=F, warning=F>>=
data <- ans.summary[,c("lineup", "vis_search", "fig_class", "card_rot", "folding")]
grid <- expand.grid(x=1:ncol(data), y=1:ncol(data))
grid <- subset(grid, x != y & x<y)

@@

ggplot()+
  geom_point(data=subset(all, points), aes(x=x, y=y)) + 
  geom_smooth(data=subset(all, points), aes(x=x, y=y), method="lm") + 
  geom_text(data=subset(all, !points), aes(x=x, y=y, label=paste0("Correlation = \n", label))) + 
  geom_line(data=densities, aes(x=x, y=y)) + 
  facet_grid(xvar~yvar, scales="free", labeller = labeller(xvar=fix.names, yvar=fix.names)) + 
  theme_bw() + 
  theme(axis.title=element_blank())
rm(list=c("data", "all", "densities"))
@
\caption{Pairwise scatterplots of test scores. Lineup scores are most highly correlated with figure classification scores, and are also highly correlated with card rotation scores. Paper folding and card rotation scores are also highly correlated.\label{fig:scatterplotmatrix}}
\end{figure}

@@

\newtext{We see that again, PC1 accounts for general visual aptitude. PC2, however, emphasizes the relationship between figure classification and Lineup task 1 (and to a lesser extent, on lineup task 2). PC3 differentiates the lineup task 3 from the visual search task; other tasks do not impact PC3 to a meaningful degree. PC4 separates lineup test 2 from the paper folding and card rotation tests and emphasizes the overlap with the visual search test. PC5 separates lineup test 2 and the figure classification test from the overlap between the other lineup tests and the visual search test. PC 6 and 7 account for \Sexpr{as.numeric(round(100*importance[6], 1))}\% and \Sexpr{as.numeric(round(100*importance[7], 1))}\% of the variance, respectively; PC6 seems to differentiate the first two lineup tasks from the overlap between the third lineup task and the figure classification task.
}
\subsection{Lineup Plot Types}
